---
 github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

AustraliaFirePredictor
<!-- badges: start -->
<!-- badges: end -->
The goal of AustraliaFirePredictor is to provide tools for predicting fire locations and analyzing fire data in Australia using machine learning models, specifically XGBoost.

Installation
You can install the development version of AustraliaFirePredictor from GitHub with:

```{r}
# install.packages("devtools")
devtools::install_github("ishancoderr/AustraliaFirePredictor")

```
## Data Explanation

The data used in this package contains the following columns:

- **Latitude**: Center of 1km fire pixel but not necessarily the actual location of the fire as one or more fires can be detected within the 1km pixel.
- **Longitude**: Center of 1km fire pixel but not necessarily the actual location of the fire as one or more fires can be detected within the 1km pixel.
- **Brightness**: Brightness temperature 21 (Kelvin). Channel 21/22 brightness temperature of the fire pixel measured in Kelvin.
- **Scan**: Along Scan pixel size. The algorithm produces 1km fire pixels, but MODIS pixels get bigger toward the edge of scan. Scan and track reflect actual pixel size.
- **Track**: Along Track pixel size. The algorithm produces 1km fire pixels, but MODIS pixels get bigger toward the edge of scan. Scan and track reflect actual pixel size.
- **Acq_date**: Acquisition Date. Date of MODIS acquisition.
- **Acq_time**: Acquisition Time. Time of acquisition/overpass of the satellite (in UTC).
- **Satellite**: Satellite. A = Aqua and T = Terra.
- **Instrument**: Instrument. Constant value for MODIS.
- **Confidence**: Confidence (0-100%). This value is based on a collection of intermediate algorithm quantities used in the detection process. It is intended to help users gauge the quality of individual hotspot/fire pixels. Confidence estimates range between 0 and 100% and are assigned to one of the three fire classes (low-confidence fire, nominal-confidence fire, or high-confidence fire).
- **Bright_t31**: Brightness temperature 31 (Kelvin). Channel 31 brightness temperature of the fire pixel measured in Kelvin.
- **Frp**: Fire Radiative Power. Depicts the pixel-integrated fire radiative power in MW (megawatts).
- **Daynight**: Day / Night. D = Daytime, N = Nighttime.

This data is used to train machine learning models to predict fire occurrences and analyze fire patterns in Australia.

Example
This is a basic example that shows you how to use the main functions in the AustraliaFirePredictor package:
```{r}
library(AustraliaFirePredictor)

# Path to the example fire data CSV file included in the package
file_path <- system.file("extdata", "fire_data.csv", package = "AustraliaFirePredictor")

# Path to save the trained model
output_model_path <- file.path(tempdir(), "firePredictorModel.xgb")

# Run the fire location check function to train the model and evaluate it
results <- fire_location_check(file_path = file_path, output_model_path = output_model_path)

# Print the RMSE of the trained model
cat("Model RMSE:", results$rmse, "\n")

# Create example input data for prediction
input_data <- data.frame(
  longitude = -119.7,
  latitude = 34.5,
  brightness = 300.0,
  frp = 12.5,
  satellite = "Aqua",
  instrument = "MODIS",
  daynight = "D",
  scan = 1.1,
  track = 1.2,
  bright_t31 = 290.0
)

# Predict fire occurrences using the trained model
predictions <- predict_fire(input_data)

# Print the predictions
print(predictions)

```



